---
title: "Figure 3 - descriptives"
author: "Sas"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/sjurriaa/Australian Institute of Marine Science/Carly Randall - Randall Lab/CAD/CAD Year1")
```

# Load libraries

```{r}
library(readxl) 
library(writexl) 
library(dplyr) 
library(openxlsx) 
library(readr) 
library(ggplot2) 
library(patchwork) 
library(cowplot) 
library(lubridate) 
library(tidyr)
library(stringr)
```

# Read in the data

```{r}
#| label: import data
#| warning: false
#| message: false
df_size <- read_excel("data/YEAR1_Survival.xlsx", 
    col_types = c("numeric", "skip", "text", 
        "skip", "skip", "numeric", "skip", 
        "text", "text", "numeric", "skip", 
        "text", "numeric", "text", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric"), na = "na")
colnames(df_size)


# Set working directory 
setwd("C:/Users/sjurriaa/Australian Institute of Marine Science/Carly Randall - Randall Lab/CAD/CAD Field Data/2022_Deployment_Saskia_YEAR1/Survival data")  


# Import data file 
data_survival <- read_excel("2022_Deployment_Saskia_MasterData.xlsx",
    sheet = "Survival", col_types = c("skip", 
        "skip", "text", "skip", "skip", "numeric", 
        "skip", "text", "text", "text", "skip", 
        "text", "text", "text", "numeric", 
        "skip", "numeric"),
    na = "NA")

colnames(data_survival)
```

# Size

Data tidying

**Create new column for size and remove empty rows with no size data**

```{r}
#| label: mutate survival
df_size <- df_size |>
  mutate(Size = (LinLength_mm * PerpLenght_mm * Height_mm)/1000,
         Size_log = log(Size + 1)) # Adding 1 to avoid log(0) 

# Remove rows with missing values
df_size$SurvDev <- as.numeric(df_size$SurvDev)
df_size <- df_size[complete.cases(df_size$Size),  ]
```

Summary data size

```{r}
summary_size <- df_size |>
  group_by(Reef, Spp, Site) |>
  summarise(
    Mean_Size = mean(Size, na.rm = TRUE),
    SD_Size = sd(Size, na.rm = TRUE),
    N = n(),
    max_Size = max(Size, na.rm = TRUE)
  ) |>
  ungroup()

print(summary_size)

summary_by_species <- summary_size |> 
  group_by(Reef, Spp) |>
  summarise(
    Mean_Size_across_sites = mean(Mean_Size, na.rm = TRUE),
    SD_Size_across_sites = sd(Mean_Size, na.rm = TRUE),
    N_Sites = n()
  )

print(summary_by_species)
```

# Survival

### Heron - restructure

For each species:

1\. Identify if there was any survival at the device level (i.e., if any tab survived, mark device as a 'survivor').

2\. Summarise the data so that each row represents one device per census (removing duplicate rows for tabs).

3\. Retain relevant metadata (site, reef, wave energy, etc.) and bind both species together into one dataset.

```{r}
# Subset Heron data
heron_surv <- data_survival |> filter (Reef == "Heron")

# first for A tenuis: filter dataset for  A_tenuis, and then add survival and total count at group level
heron_Aten <- heron_surv |> 
  filter(Spp == "A. kenti") |>
  group_by(Census, ReefDev) |>
  mutate(SurvDev_Spp = ifelse(any(SurvTab == 1), 1, 0),
         Total_Spp = n()) |>
  ungroup()

# second: take heron_Aten dataset and make summary for each Census and ReefDev combination, retaining the first occurrence of several variables (thus remove tab2 and tab3 rows) and calculate if there was a survivor on any tab (max(SurvDev_Spp)).
heron_Aten2 <- heron_Aten |>
  group_by(Census, ReefDev) |>
  summarise(ExpDay = first(ExpDay),
            Reef = first(Reef),
            Site = first(Site),
            WaveEnergyLevel = first(WaveEnergyLevel),
            Spp = first(Spp),
            Total_Spp = first(Total_Spp),
            SurvDev_Spp = max(SurvDev_Spp)) |>
  ungroup()

# now do same for Ahyacinthus
heron_Ahya <- heron_surv |> 
  filter(Spp == "A. hyacinthus") |>
  group_by(Census, ReefDev) |>
  mutate(SurvDev_Spp = ifelse(any(SurvTab == 1), 1, 0),
         Total_Spp = n()) |>
  ungroup()

heron_Ahya2 <- heron_Ahya |>
  group_by(Census, ReefDev) |>
  summarise(ExpDay = first(ExpDay),
            Reef = first(Reef),
            Site = first(Site),
            WaveEnergyLevel = first(WaveEnergyLevel),
            Spp = first(Spp),
            Total_Spp = first(Total_Spp),
            SurvDev_Spp = max(SurvDev_Spp)) |>
  ungroup()

#bind them together to create one dataframe
heron_device <- rbind(heron_Ahya2, heron_Aten2)
heron_device <- heron_device |>
  arrange(Census, Site, Spp, ReefDev)

#free up space
remove(heron_Aten, heron_Aten2, heron_Ahya, heron_Ahya2)

colnames(heron_device)
```

### Summary: Percent Device Survival 

```{r}
summary_heron_device <- heron_device |>
  filter(!is.na(SurvDev_Spp)) |>
  group_by(Reef, Site, WaveEnergyLevel, Spp, Census, ExpDay) |>
  summarise(SurvivalNDevices = sum(SurvDev_Spp, na.rm=T),
            DevTot =25,
            Perc_SurvDev = sum(SurvDev_Spp, na.rm =T)/25)

colnames(summary_heron_device)

summary_daviesmoore_device <- data_survival |>
  filter(!is.na(SurvDev),
         !Reef == "Heron") |>
  group_by(Reef, Site, WaveEnergyLevel, Census, ExpDay, Spp) |>
  summarise(SurvivalNDevices = sum(SurvDev, na.rm=T),
            DevTot =25,
            Perc_SurvDev = sum(SurvDev, na.rm =T)/25)

colnames(summary_daviesmoore_device)

#bind dataframes together
summary_survival_device <- rbind(summary_heron_device, summary_daviesmoore_device)
colnames(summary_survival_device)

remove(summary_daviesmoore_device, summary_heron_device, heron_device)
```

### Summary: Percent Tab Survival 

```{r}
summary_daviesmoore_tab <- data_survival |>
   filter(!Reef == "Heron") |>
  group_by(Reef, Site, WaveEnergyLevel, Census, ExpDay, Spp) |>
  summarise(SurvivalNTab = sum(SurvTab, na.rm=T),
            TabTot =75,
            Perc_SurvTab = sum(SurvDev, na.rm =T)/75)
colnames(summary_daviesmoore_tab)

summary_heron_tab <- data_survival |>
  filter(Reef == "Heron") |>
   group_by(Reef, Site, WaveEnergyLevel, Census, ExpDay, Spp) |>
   summarise(SurvivalNTabs = (sum(SurvTab, na.rm = TRUE))) 

summary_heron37 <- summary_heron_tab |>
  filter(Site %in% c("H1A", "H2B", "H3A", "H4B", "H5A"))|>
    mutate(TabTot = rep(c(37,38)),
         Perc_SurvTab = SurvivalNTabs/TabTot)

summary_heron38 <- summary_heron_tab |>
  filter(Reef == "Heron") |>
  filter(Site %in% c("H1B", "H2A", "H3B", "H4A", "H5B")) |>
    mutate(TabTot = rep(c(38, 37)),
         Perc_SurvTab = SurvivalNTabs/TabTot)

summary_heron_tab <- rbind(summary_heron37,summary_heron38)
colnames(summary_heron_tab)

summary_survival_tab <- rbind(summary_daviesmoore_tab, summary_heron_tab)

remove(summary_heron37, summary_heron38, summary_daviesmoore_tab, summary_heron_tab)
```

```{r}
moore_final_dev <- summary_survival_device |> filter(Reef == "Moore", Census =="t6")
mean(moore_final_dev$Perc_SurvDev)
sd(moore_final_dev$Perc_SurvDev)
min(moore_final_dev$Perc_SurvDev)
max(moore_final_dev$Perc_SurvDev)
```

```{r}
moore_final_tab <- summary_survival_tab |> filter(Reef == "Moore", Census =="t6")
mean(moore_final_tab$Perc_SurvTab)
sd(moore_final_tab$Perc_SurvTab)
min(moore_final_tab$Perc_SurvTab)
max(moore_final_tab$Perc_SurvTab)
```

```{r}
davies_final_dev <- summary_survival_device |> filter(Reef == "Davies", Census =="t6")
mean(davies_final_dev$Perc_SurvDev)
sd(davies_final_dev$Perc_SurvDev)
min(davies_final_dev$Perc_SurvDev)
max(davies_final_dev$Perc_SurvDev)
```

```{r}
davies_final_tab <- summary_survival_tab |> filter(Reef == "Davies", Census =="t6")
mean(davies_final_tab$Perc_SurvTab)
sd(davies_final_tab$Perc_SurvTab)
min(davies_final_tab$Perc_SurvTab)
max(davies_final_tab$Perc_SurvTab)
```

```{r}
heron_final_dev_kenti <- summary_survival_device |> filter(Reef == "Heron", Census == "t3", Spp == "A. kenti")
mean(heron_final_dev_kenti$Perc_SurvDev)
sd(heron_final_dev_kenti$Perc_SurvDev)
min(heron_final_dev_kenti$Perc_SurvDev)
max(heron_final_dev_kenti$Perc_SurvDev)
```

```{r}
heron_final_dev_hya <- summary_survival_device |> filter(Reef == "Heron", Census == "t3", Spp == "A. hyacinthus")
mean(heron_final_dev_hya$Perc_SurvDev)
sd(heron_final_dev_hya$Perc_SurvDev)
min(heron_final_dev_hya$Perc_SurvDev)
max(heron_final_dev_hya$Perc_SurvDev)
```

```{r}
heron_final_tab_kenti <- summary_survival_tab |> filter(Reef == "Heron", Census =="t3", Spp == "A. kenti")
mean(heron_final_tab_kenti$Perc_SurvTab)
sd(heron_final_tab_kenti$Perc_SurvTab)
min(heron_final_tab_kenti$Perc_SurvTab)
max(heron_final_tab_kenti$Perc_SurvTab)
```

```{r}
heron_final_tab_hya <- summary_survival_tab |> filter(Reef == "Heron", Census =="t3", Spp == "A. hyacinthus")
mean(heron_final_tab_hya$Perc_SurvTab)
sd(heron_final_tab_hya$Perc_SurvTab)
min(heron_final_tab_hya$Perc_SurvTab)
max(heron_final_tab_hya$Perc_SurvTab)
```

# Figures

### Figure survival

```{r}
# Calculate average survival and standard error for each species within each reef at each time point
summary_survival_stats <- summary_survival_device |>
  group_by(Reef, Census, Spp) |>
  summarise(Avg_Survival = mean(Perc_SurvDev, na.rm = TRUE),
            SE = sd(Perc_SurvDev, na.rm = TRUE) / sqrt(n())) |>
    ungroup() |>
  mutate(ExpDay = c(0, 103, 163, 255, 326, 527, 0, 0, 71, 71, 415, 415, 843,843,0, 92,151,247, 337, 525))

# Reorder the levels of the Reef variable
summary_survival_stats$Reef <- factor(summary_survival_stats$Reef, levels = c("Moore", "Davies", "Heron"))

# Plot the average survival over time for each species with standard error bars
plot_survival_time <- ggplot(summary_survival_stats, aes(x = ExpDay, y = Avg_Survival, color = Spp)) +
   geom_point(size=3) +
  scale_color_manual(values = c("A. millepora" = "green", "A. hyacinthus" = "orange", "A. kenti" = "purple")) +
  geom_errorbar(aes(ymin = Avg_Survival - SE, ymax = Avg_Survival + SE), width = 0.2) +
  facet_wrap(~Reef, scales = "free") +
  labs(x = "Experiment Day",
       y = "% Survival (yield)",
       color = "Species") +
  theme_bw() +
  ylim(0,1) +
  xlim(0,900) +
   theme(panel.border = element_rect(color = "black", fill = NA),
        axis.line = element_line(color = "black"),
        axis.text = element_text(size = 13),
        axis.title = element_text(size = 14), 
        legend.text = element_text(size = 13),
        strip.text = element_blank())

plot_survival_time


```

### Boxplot size

```{r}
# Reorder the levels of the Reef variable
df_size$Reef <- factor(df_size$Reef, levels = c("Moore", "Davies", "Heron"))

# Update the Site column in the size_allreef dataset
df_size$Site <- gsub("A$", "a", df_size$Site)
df_size$Site <- gsub("B$", "b", df_size$Site)
unique(df_size$Site)

# Cap the y-values for Heron only. This keeps the data but limits extreme values visually.
df_size <- df_size |>
  mutate(Size_capped = ifelse(Reef == "Heron" & Size > 200, 200, Size))


# Plot the average growth for each reef and species - BOXPLOT
plot_size_site <- ggplot(df_size, aes(x= Site, y = Size_capped, color = Spp)) +
   geom_boxplot() +
  scale_color_manual(values = c("A. millepora" = "green", "A. hyacinthus" = "orange", "A. kenti" = "purple")) +
  facet_wrap(~Reef, scales = "free") +
  labs(x = "Site",
       y = "Coral Size (cm3)",
       color = "Species") +
  theme_bw() +
  theme(panel.border = element_rect(color = "black", fill = NA),
        axis.line = element_line(color = "black"),
        axis.text = element_text(size = 13),
        axis.title = element_text(size = 14), 
        legend.text = element_text(size = 13),
        strip.text = element_blank())

plot_size_site
```

# Multiplot

```{r}
library(gridExtra)
library(gtable)
library(grid)
library(egg)

grid_arrange_shared_legend <-
  function(...,
           ncol = length(list(...)),
           nrow = 1,
           position = c("bottom", "right")) {
    
    plots <- list(...)
    position <- match.arg(position)
    g <-
      ggplotGrob(plots[[1]] + theme(legend.position = position))$grobs
    legend <- g[[which(sapply(g, function(x)
      x$name) == "guide-box")]]
    lheight <- sum(legend$height)
    lwidth <- sum(legend$width)
    gl <- lapply(plots, function(x)
      x + theme(legend.position = "none"))
    gl <- c(gl, ncol = ncol, nrow = nrow)
    
    combined <- switch(
      position,
      "bottom" = arrangeGrob(
        do.call(arrangeGrob, gl),
        legend,
        ncol = 1,
        heights = unit.c(unit(1, "npc") - lheight, lheight)
      ),
      "right" = arrangeGrob(
        do.call(arrangeGrob, gl),
        legend,
        ncol = 2,
        widths = unit.c(unit(1, "npc") - lwidth, lwidth)
      )
    )
    
    grid.newpage()
    grid.draw(combined)
    
    # return gtable invisibly
    invisible(combined)
    
  }
```

```{r}
capture_plot <- function() {
  grid_arrange_shared_legend(
    plot_survival_time,plot_size_site,
    ncol = 1, 
    nrow = 2)
}

capture_plot()
```

```{r}

# Save the plot using ggsave
ggsave("Figure descriptives_heron_capped.jpeg", plot = capture_plot(), width = 13, height = 6)

```
