---
title: "Survival_summary"
author: "Saskia"
format: html
editor: visual
---

# Data preparation

### Load the required packages

```{r}
library(readxl) 
library(writexl) 
library(dplyr) 
library(openxlsx) 
library(readr) 
library(ggplot2) 
library(patchwork) 
library(cowplot) 
library(lubridate) 
library(tidyr)
library(stringr)
```

### Set the workdrive and load data

```{r}
# Set working directory 
setwd("C:/Users/sjurriaa/Australian Institute of Marine Science/Carly Randall - Randall Lab/CAD/CAD Field Data/2022_Deployment_Saskia_YEAR1/Survival data")  


# Import data file 
data_survival <- read_excel("2022_Deployment_Saskia_MasterData.xlsx",
    sheet = "Survival", col_types = c("skip", 
        "skip", "text", "skip", "skip", "numeric", 
        "skip", "text", "text", "text", "skip", 
        "text", "text", "text", "numeric", 
        "skip", "numeric"),
    na = "NA")
```

### \-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\--

# Data summary

### First HERON: summary using species at device level

```{r}
data_survival <- data_survival |>
  select(-c(Observer, CensusT, DeploymentDate, CensusDate, Device_ID, NoRecruits))
heron_surv <- data_survival |> filter (Reef == "Heron")

# first for A tenuis: filter dataset for  A_tenuis, and then add survival and total count at group level
heron_Aten <- heron_surv |> 
  filter(Spp == "A. kenti") |>
  group_by(Census, ReefDev) |>
  mutate(SurvDev_Spp = ifelse(any(SurvTab == 1), 1, 0),
         Total_Spp = n()) |>
  ungroup()

# second: take heron_Aten dataset and make summary for each Census and ReefDev combination, retaining the first occurrence of several variables (thus remove tab2 and tab3 rows) and calculate if there was a survivor on any tab (max(SurvDev_Spp)).
heron_Aten2 <- heron_Aten |>
  group_by(Census, ReefDev) |>
  summarise(ExpDay = first(ExpDay),
            Reef = first(Reef),
            Site = first(Site),
            WaveEnergyLevel = first(WaveEnergyLevel),
            Spp = first(Spp),
            Total_Spp = first(Total_Spp),
            SurvDev_Spp = max(SurvDev_Spp)) |>
  ungroup()

# now do same for Ahyacinthus
heron_Ahya <- heron_surv |> 
  filter(Spp == "A. hyacinthus") |>
  group_by(Census, ReefDev) |>
  mutate(SurvDev_Spp = ifelse(any(SurvTab == 1), 1, 0),
         Total_Spp = n()) |>
  ungroup()

heron_Ahya2 <- heron_Ahya |>
  group_by(Census, ReefDev) |>
  summarise(ExpDay = first(ExpDay),
            Reef = first(Reef),
            Site = first(Site),
            WaveEnergyLevel = first(WaveEnergyLevel),
            Spp = first(Spp),
            Total_Spp = first(Total_Spp),
            SurvDev_Spp = max(SurvDev_Spp)) |>
  ungroup()

#bind them together to create one dataframe
heron_Spp <- rbind(heron_Ahya2, heron_Aten2)
heron_Spp <- heron_Spp |>
  arrange(Census, Site, Spp, ReefDev)

#free up space
remove(heron_Aten, heron_Aten2, heron_Ahya, heron_Ahya2)
```

### Create dataset with % survival Devices

```{r}
summary_heron <- heron_Spp |>
  filter(!is.na(SurvDev_Spp)) |>
  group_by(Reef, Site, WaveEnergyLevel, Spp, Census, ExpDay) |>
  summarise(SurvivalNDevices = sum(SurvDev_Spp, na.rm=T),
            DevTot =25,
            Perc_SurvDev = sum(SurvDev_Spp, na.rm =T)/25)
```

### Create dataset for final timepoint Heron Spp

```{r}
heron_Spp_t3 <- heron_Spp |>
  filter(Census == "t3")

summary_heront3 <- heron_Spp_t3 |>
  filter(!is.na(SurvDev_Spp)) |>
  group_by(Reef, Site, WaveEnergyLevel, Spp, Census, ExpDay) |>
  summarise(SurvivalNDevices = sum(SurvDev_Spp, na.rm=T),
            DevTot =25,
            Perc_SurvDev = sum(SurvDev_Spp, na.rm =T)/25)
```

### Summary % survival - dataframes LONG

Create a summary survival dataframe with a new column that sums up the survival score of all the tabs for each site for each census. Then create an additional dataframe with a new column that sums up the survival score of the devices for each site for each census. Last, combine both dataframes so that there it shows for each site and per census a column with the total number of tabs with survivorship, and the total number of devices with at least one survivor.

```{r}
summary_survival_Dev <- data_survival |>
  filter(!is.na(SurvDev)) |>
  group_by(Reef, Site, WaveEnergyLevel, Census, ExpDay, Spp) |>
  summarise(SurvivalNDevices = sum(SurvDev, na.rm=T),
            DevTot =25,
            Perc_SurvDev = sum(SurvDev, na.rm =T)/25)
```

### Combine Heron (yield calculated by Spp) with Moore and Davies (Spp)

```{r}
summary_daviesmoore <- summary_survival_Dev |>
  filter(!Reef == "Heron")

#heron dataframe: summary_heron
#moore davies dataframe: summary_survival_Dev

#bind dataframes together
summary_survival <- rbind(summary_heron, summary_daviesmoore)
```

Create summary dataframe that includes species.

```{r}
summary_survival_Spp <- data_survival |>
  group_by(Reef, Site, WaveEnergyLevel, Census, ExpDay, Spp) |>
  summarise(SurvivalNTabs = (sum(SurvTab, na.rm = TRUE))) 

summary_survival_Spp_DavMoore <- summary_survival_Spp |>
  filter(!Reef == "Heron") |>
  mutate(TabTot = 75,
         Perc_SurvTab = SurvivalNTabs/TabTot)

summary_survival_Spp_Heron37 <- summary_survival_Spp |>
  filter(Reef == "Heron") |>
  filter(Site %in% c("H1A", "H2B", "H3A", "H4B", "H5A"))|>
    mutate(TabTot = rep(c(37,38)),
         Perc_SurvTab = SurvivalNTabs/TabTot)

summary_survival_Spp_Heron38 <- summary_survival_Spp |>
  filter(Reef == "Heron") |>
  filter(Site %in% c("H1B", "H2A", "H3B", "H4A", "H5B")) |>
    mutate(TabTot = rep(c(38, 37)),
         Perc_SurvTab = SurvivalNTabs/TabTot)

summary_survival_Spp_Heron <- rbind(summary_survival_Spp_Heron37,summary_survival_Spp_Heron38)

summary_survival_Spp <- rbind(summary_survival_Spp_DavMoore, summary_survival_Spp_Heron)

remove(summary_survival_Spp_Heron, summary_survival_Spp_DavMoore, summary_survival_Spp_Heron37, summary_survival_Spp_Heron38)
```

```{r}
summary_survival <- summary_survival_Dev |>
  left_join(
    summary_survival_Spp)

remove(summary_survival_Dev, summary_survival_Spp)
```

[**Filter to last census (t3 for Heron, t6 for Davies and Moore)**]{.underline}

```{r}
summary_survival_t6 <- summary_survival |>   
  filter(Census == "t6")           #Only show survival at the last timepoint

summary_survival_t2 <- summary_survival |>   
  filter(Census == "t3") |>
  filter(Reef == "Heron")         #only show survival for Heron at t3

summary_survival_finalcensus <- rbind(summary_survival_t6, summary_survival_t2)

remove(summary_survival_t2, summary_survival_t6)
```

[**Overall mean % survival by site**]{.underline}

#NOT RELEVANT, only % at t6 is relevant!! - Get the overall mean percentage survival per site across timepoints.

```{r}
survival_mean_Tab <- survival_perc_Tab |>
  rowwise() |>
  mutate(MeanPerc = mean(c_across(starts_with("Perc")), na.rm = TRUE),
         SDPerc = sd(c_across(starts_with("Perc")), na.rm = TRUE),
         SEPerc = SDPerc/sqrt((NTabs)))
```

```{r}
survival_mean_dev <- survival_perc_dev |>
  rowwise() |>
  mutate(MeanPerc = mean(c_across(starts_with("Perc")), na.rm = TRUE),
         SDPerc = sd(c_across(starts_with("Perc")), na.rm = TRUE),
         SEPerc = SDPerc/sqrt((NDev)))
```

Merge summary_heron (with device level percentage survival for species) with summary moore and davies with device level percentage survival

```{r}
#remove heron from the main dataframe because the device level spp percentages are not accurate
summary_final <- summary_survival_finalcensus |>
  filter(Reef %in% c("Davies", "Moore"))

#make WaveEnergyLevel a character in the heron dataframe
summary_heron$WaveEnergyLevel <- as.character(summary_heron$WaveEnergyLevel)

#bind both summaries together
summary_final <- summary_final |>
  rbind(summary_heron)
```

### Fig1a - Boxplot - Mean survival Reef-tabs

```{r, fig.show='all', echo=FALSE}
# Plot 1a - Reef (tab)
plot1a <- ggplot(summary_survival, aes(x = Reef, y = Perc_SurvTab, color = Reef)) +
  geom_boxplot() +
  labs(title = "Mean Survival (tabs)",
       x = "Reef",
       y = "% Survival") +
  theme_minimal() +
  theme(panel.border = element_rect(color = "black", fill = NA),
        axis.line = element_line(color = "black"),
        panel.grid.major = element_line(color = "gray", linetype = "dashed"),
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 12),  # Adjust font size of axis text
        axis.title = element_text(size = 14),
        legend.text = element_text(size = 12))  # Adjust font size of axis titles)

plot1a

ggsave("Fig1a - Mean survival Reef-tabs.png", plot=plot1a, width = 8, height = 6, dpi = 300)
```

### Fig1b - Boxplot - Mean survival Reef-yield

```{r}
plot1b <- ggplot(summary_survival, aes(x = Reef, y = Perc_SurvDev, color = Reef)) +   
  geom_boxplot() +   
  labs(title = "Mean Survival (yield)", x = "Reef", y = "% Survival") +   
  theme_minimal() +   
  theme(panel.border = element_rect(color = "black", fill = NA),         
        axis.line = element_line(color = "black"),         
        panel.grid.major = element_line(color = "gray",linetype="dashed"),
        panel.grid.minor = element_blank(),         
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14),         
        legend.text = element_text(size = 12))
plot1b 

ggsave("Fig1b - Mean survival Reef-yield.png", plot=plot1b, width = 8, height = 6, dpi = 300) 
```

### 

### Fig2a - Boxplot - Mean survival final census Reef-tabs

```{r}
# Plot 2a - Reef (tab)
plot2a <- ggplot(summary_survival_finalcensus, aes(x = Reef, y = Perc_SurvTab, color = Reef)) +
  geom_boxplot() +
  labs(title = "Mean survival final census (tabs)",
       x = "Reef",
       y = "% Survival") +
  theme_minimal() +
  theme(panel.border = element_rect(color = "black", fill = NA),
        axis.line = element_line(color = "black"),
        panel.grid.major = element_line(color = "gray", linetype = "dashed"),
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 12),  # Adjust font size of axis text
        axis.title = element_text(size = 14),
        legend.text = element_text(size = 12))  # Adjust font size of axis titles)

plot2a

ggsave("Fig2a - Mean survival final census Reef-tabs.png", plot=plot2a, width = 8, height = 6, dpi = 300)
```

### Fig2b - Boxplot - Mean survival final census Reef-yield

```{r}
# Reorder the levels of the Reef variable
summary_survival_finalcensus$Reef <- factor(summary_survival_finalcensus$Reef, levels = c("Moore", "Davies", "Heron"))

plot2b <- ggplot(summary_survival_finalcensus, aes(x = Reef, y = Perc_SurvDev, colour = Reef)) +
  geom_boxplot() +
  labs(title = "Mean Survival final census (yield)",
       x = "Reef",
       y = "% Survival") +
  theme_minimal() +
  theme(panel.border = element_rect(color = "black", fill = NA),
        axis.line = element_line(color = "black"),
        panel.grid.major = element_line(color = "gray", linetype = "dashed"),
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 12),  # Adjust font size of axis text
        axis.title = element_text(size = 14))

plot2b
ggsave("Fig2b - Mean survival final census Reef-yield.jpeg", plot=plot2b, width = 8, height = 6, dpi = 300)

```

### Fig2c - Boxplot - Mean survival final census Reef-yield per site and species

```{r}
library(cowplot)

final_moore <- summary_final |>
  filter(Reef == "Moore")
final_davies <- summary_final |>
  filter(Reef == "Davies")
final_heron <- summary_final |>
  filter(Reef == "Heron")


plot_moore <- ggplot(final_moore, aes(x = Site, y = Perc_SurvDev, colour = Spp)) +
  geom_boxplot() +
  labs(title = "Moore Reef",
       x = "Site",
       y = "% Survival (yield)") +
  theme_minimal() +
  scale_colour_manual(values = c("green")) +
  theme(panel.border = element_rect(color = "black", fill = NA),
        axis.line = element_line(color = "black"),
        panel.grid.major = element_line(color = "gray", linetype = "dashed"),
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 12),  # Adjust font size of axis text
        axis.title = element_text(size = 14))+
    theme(legend.position = "bottom") +
  scale_y_continuous(limits = c(0, 0.8))  # Adjust y-axis range


plot_davies <- ggplot(final_davies, aes(x = Site, y = Perc_SurvDev, colour = Spp)) +
  geom_boxplot() +
  labs(title = "Davies Reef",
       x = "Site",
       y = NULL) +
  theme_minimal() +
  scale_colour_manual(values = c("orange")) +
  theme(panel.border = element_rect(color = "black", fill = NA),
        axis.line = element_line(color = "black"),
        panel.grid.major = element_line(color = "gray", linetype = "dashed"),
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 12),  # Adjust font size of axis text
        axis.title = element_text(size = 14))+
    theme(legend.position = "bottom") +
  scale_y_continuous(limits = c(0, 0.8))  # Adjust y-axis range


plot_heron <- ggplot(final_heron, aes(x = Site, y = Perc_SurvDev, colour = Spp)) +
  geom_boxplot() +
  labs(title = "Heron Reef",
       x = "Site",
       y= NULL) +
  theme_minimal() +
  scale_colour_manual(values = c("orange", "purple")) + 
  theme(panel.border = element_rect(color = "black", fill = NA),
        axis.line = element_line(color = "black"),
        panel.grid.major = element_line(color = "gray", linetype = "dashed"),
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 12),  # Adjust font size of axis text
        axis.title = element_text(size = 14)) +
    theme(legend.position = "bottom") +
  scale_y_continuous(limits = c(0, 0.8))  # Adjust y-axis range


plot_moore
plot_davies
plot_heron

ggsave("Fig2c_moore - Moore final census - yield.png", plot=plot_moore, width = 4, height = 6, dpi = 300)
ggsave("Fig2c_davies - Davies final census - yield.png", plot=plot_davies, width = 4, height = 6, dpi = 300)
ggsave("Fig2c_heron - Heron final census - yield.png", plot=plot_heron, width = 4, height = 6, dpi = 300)

```

### Fig3 - Point - % survival final census yield over sites

```{r}
# Create the plot
plot3 <- ggplot(summary_survival_finalcensus, aes(x = Site, y = Perc_SurvDev, color = Reef)) +
  geom_point(size=3) +
  labs(title = "Survival by sites (final census)",
       x = "Sites",
       y = "% Survival (device yield)") +
  theme_bw() +
  ylim(0,1) +
  theme(panel.border = element_rect(color = "black", fill = NA),
        axis.line = element_line(color = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14), 
        legend.text = element_text(size = 12)) +
  facet_wrap(~Reef, scales = "free")

plot3
ggsave("Fig3 - Survival final census - yield.png", plot=plot3, width = 14, height = 6, dpi = 300)

```

### Fig4 - Point - % survival Species

```{r}

plot4<- ggplot(summary_survival_finalcensus, aes(x = Site, y = Perc_SurvTab, color = Reef)) +
  geom_point(size=3) +  
  labs(title = "Survival final census",
       x = "Sites",
       y = "% Survival (tabs)") +
  theme_bw() +
  ylim(-0.01,1) +
  theme(panel.border = element_rect(color = "black", fill = NA),
        axis.line = element_line(color = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14), 
        legend.text = element_text(size = 12)) +
  facet_wrap(~Spp, scales = "free")
plot4

ggsave("Fig4 - Survival by species - tabs.png", plot=plot4, width = 14, height = 6, dpi = 300)

```

### Fig5 - Point - % survival final census yield over ExpDay

```{r}
# Plot with specific days on the x-axis for each reef
plot5 <- ggplot(summary_survival, aes(x = ExpDay, y = Perc_SurvDev, color = Reef)) +
  geom_point(size=3) +
  labs(title = "Survival by time",
       x = "Experiment days",
       y = "% Survival (device yield)") +
  theme_bw() +
  ylim(0,1) +
   xlim(0,900) +
  theme(panel.border = element_rect(color = "black", fill = NA),
        axis.line = element_line(color = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14), 
        legend.text = element_text(size = 12)) +
  facet_wrap(~Reef, scales = "free")

plot5
ggsave("Fig5 - Survival over time - yield.png", plot=plot5, width = 14, height = 6, dpi = 300)

```

### Fig6- Point - % survival Species over ExpDay

```{r}
plot6 <- ggplot(summary_survival, aes(x = ExpDay, y = Perc_SurvTab, color = Reef)) +
  geom_point(size=3) +
  labs(title = "Survival by species",
       x = "Experiment days",
       y = "% Survival (tabs)") +
  theme_bw() +
  ylim(0,1) +
  xlim(0,900) +
  theme(panel.border = element_rect(color = "black", fill = NA),
        axis.line = element_line(color = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14), 
        legend.text = element_text(size = 12)) +
  facet_wrap(~Spp, scales = "free")

plot6
ggsave("Fig6 - Survival Spp over time - tab.png", plot=plot6, width = 14, height = 6, dpi = 300)

```

### Fig7a- Point - Mean % survival Species (average sites) over ExpDay

```{r}
# Calculate average survival and standard error for each species within each reef at each time point
summary_survival_stats <- summary_survival |>
  group_by(Reef, Census, Spp) |>
  summarise(Avg_Survival = mean(Perc_SurvDev, na.rm = TRUE),
            SE = sd(Perc_SurvDev, na.rm = TRUE) / sqrt(n())) |>
    ungroup() |>
  mutate(ExpDay = c(0, 103, 163, 255, 326, 527, 0, 0, 71, 71, 415, 415, 843,843,0, 92,151,247, 337, 525))

# Reorder the levels of the Reef variable
summary_survival_stats$Reef <- factor(summary_survival_stats$Reef, levels = c("Moore", "Davies", "Heron"))

# Plot the average survival over time for each species with standard error bars
plot7a <- ggplot(summary_survival_stats, aes(x = ExpDay, y = Avg_Survival, color = Spp)) +
   geom_point(size=3) +
  scale_color_manual(values = c("A. millepora" = "green", "A. hyacinthus" = "orange", "A. kenti" = "purple")) +
  geom_errorbar(aes(ymin = Avg_Survival - SE, ymax = Avg_Survival + SE), width = 0.2) +
  facet_wrap(~Reef, scales = "free") +
  labs(title = "Mean Survival Over Time by Species and Reef",
       x = "Experiment Day",
       y = "% Survival (yield)",
       color = "Species") +
  theme_bw() +
  ylim(0,1) +
  xlim(0,900) +
  theme(panel.border = element_rect(color = "black", fill = NA),
        axis.line = element_line(color = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14), 
        legend.text = element_text(size = 12))

plot7a
ggsave("Fig7a - Mean survival Spp over time - tab.png", plot=plot7a, width = 14, height = 6, dpi = 300)

```

### Fig7b- Point - % survival Species over sites at FINAL TIMEPOINTS

```{r}
final_survival1 <- summary_survival %>%
  filter(Census == "t6")
final_survival2 <- summary_survival %>%
  filter(Reef == "Heron",
         Census =="t3")

final_survival <- rbind(final_survival1, final_survival2)

remove(final_survival1, final_survival2)

# Reorder the levels of the Reef variable
final_survival$Reef <- factor(final_survival$Reef, levels = c("Moore", "Davies", "Heron"))

plot7b <- ggplot(final_survival, aes(x = Site, y = Perc_SurvDev, color = Spp))  +
   geom_point(size=3) +
  scale_color_manual(values = c("A. millepora" = "green", "A. hyacinthus" = "orange", "A. kenti" = "purple")) +
  facet_wrap(~Reef, scales = "free") +
  labs(title = "Final Survival of Species by Site",
       x = "Site",
       y = "% Survival (yield)",
       color = "Species") +
  theme_bw() +
  theme(panel.border = element_rect(color = "black", fill = NA),
        axis.line = element_line(color = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14), 
        legend.text = element_text(size = 12)) +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01),  limits = c(0, 0.80))


# Display plot
plot7b
ggsave("Fig7b - Survival Spp over site FINAL TIMEPOINT - tab.png", plot=plot7b, width = 14, height = 6, dpi = 300)



```

### Fig8- Point - Mean % survival Species grouped by Reef (average sites) over ExpDay

```{r}
# Calculate average survival and standard error for each species within each reef at each time point

summary_survival_stats <- summary_survival |>
  group_by(Reef, Census, Spp) |>
  summarise(Avg_Survival = mean(Perc_SurvTab, na.rm = TRUE),
            SE = sd(Perc_SurvTab, na.rm = TRUE) / sqrt(n())) 

|>
    ungroup() |>
  mutate(ExpDay = c(0, 103, 163, 255, 326, 527, 0, 0, 71, 71, 415, 415, 843,843,0, 92,151,247, 337, 525))

# Plot the average survival over time for each species with standard error bars
plot8 <- ggplot(summary_survival_stats, aes(x = ExpDay, y = Avg_Survival, color = Spp)) +
  geom_point(size=3) +
  geom_errorbar(aes(ymin = Avg_Survival - SE, ymax = Avg_Survival + SE), width = 0.2) +
  facet_wrap(~Reef, scales = "free") +
  labs(title = "Mean Survival Over Time by Species and Reef",
       x = "Experiment Day",
       y = "Average % Survival (tabs)",
       color = "Species") +
  theme_bw() +
  ylim(0,1) +
  xlim(0,900) +
  theme(panel.border = element_rect(color = "black", fill = NA),
        axis.line = element_line(color = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14), 
        legend.text = element_text(size = 12))

plot8
ggsave("Fig8 - Mean survival Spp by Reef over time - tab.png", plot=plot8, width = 14, height = 6, dpi = 300)
```

### Save datasets

```{r}
# Specify the desired file name and path
file_name <- "Summary Survival YEAR1.xlsx"
file_path <- paste0("C:/Users/sjurriaa/Australian Institute of Marine Science/Carly Randall - Randall Lab/CAD/CAD Year1/output/", file_name)

# Write the dataset to an Excel file
write.xlsx(summary_survival, file_path)


# Specify the desired file name and path
file_name <- "Summary Survival YEAR1 FinalCensus.xlsx"
file_path <- paste0("C:/Users/sjurriaa/Australian Institute of Marine Science/Carly Randall - Randall Lab/CAD/CAD Year1/output/", file_name)

# Write the dataset to an Excel file
write.xlsx(summary_survival_finalcensus, file_path)
```

### \-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\--

# Plot data against current meters

### Load the data

```{r}
data_env <- read_excel("C:/Users/sjurriaa/Australian Institute of Marine Science/Carly Randall - Randall Lab/CAD/CAD Field Data/2022_Deployment_Saskia_YEAR1/Instruments data/Marotte_HS_all_reefs/Metrics selection/FinalMetrics_Summary.xlsx", 
    col_types = c("text", "text",  
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "text", "text", "text", "text"), 
    na = "na")

# free up space
remove(plot1a, plot1b, plot2a,plot2b, plot3, plot4, plot5, plot6, plot7, plot8)
```

### Merge dataframes

Merge summary data with environmental data

```{r}
merged_data <- merge(summary_survival, data_env, by = c("Reef", "Site"), all.x = TRUE)
```

Make selection for final census

```{r}
merged_data_t6 <- merged_data |>   
  filter(Census == "t6")           #Only show survival at the last timepoint

merged_data_t2 <- merged_data |>   
  filter(Census == "t3") |>
  filter(Reef == "Heron")         #only show survival for Heron at t3

merged_data_finalcensus <- rbind(merged_data_t6, merged_data_t2)

remove(merged_data_t2, merged_data_t6)

```

# Figures

### Fig9a - Point - survival final census over nominal Wave Energy Level

```{r}
# Convert WaveEnergyLevel column to numeric
merged_data_finalcensus$WaveEnergyLevel <- as.numeric(merged_data_finalcensus$WaveEnergyLevel)


plot9 <- ggplot(merged_data_finalcensus, aes(x = WaveEnergyLevel, y = Perc_SurvDev, color = Reef)) +
  geom_point(size=3) +
  geom_smooth() + 
  labs(title = "Survival over nominal Wave Energy Level (final census)",
       x = "nominal Wave Energy Level",
       y = "% Survival (yield)") +
  theme_bw() +
  ylim(0,1) +
  theme(panel.border = element_rect(color = "black", fill = NA),
        axis.line = element_line(color = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14), 
        legend.text = element_text(size = 12)) +
  facet_wrap(~Reef, scales = "free")

plot9
ggsave("Fig9 - Survival final census WaveEnergyLevel - yield.png", plot=plot9, width = 14, height = 6, dpi = 300)
```

### Fig9b - Point - survival over time by WaveEnergyLevel and Reef

```{r}
# Reorder the levels of the Reef variable
merged_data$Reef <- factor(merged_data$Reef, levels = c("Moore", "Davies", "Heron"))

plot9b <- ggplot(merged_data, aes(x = ExpDay, y = Perc_SurvTab, color = Spp)) +
  geom_point(size=3) +
  labs(title = "Survival across nominal Wave Energy Level",
       x = "Experiment days",
       y = "% Survival (tabs)") +
  geom_smooth(method="lm")+
  theme_bw() +
  ylim(0,1) +
  theme(panel.border = element_rect(color = "black", fill = NA),
        axis.line = element_line(color = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14), 
        legend.text = element_text(size = 12)) +
  facet_grid(Reef~ WaveEnergyLevel, scales = "free", space = "free")

plot9b
ggsave("Fig9b - Survival final census WaveEnergyLevel - yield.png", plot=plot9b, width = 14, height = 6, dpi = 600)
```

### Fig10 - Point - survival final census over Ub

```{r}
plot10 <- ggplot(merged_data_finalcensus, aes(x = Ub_avrg, y = Perc_SurvDev, color = Reef)) +
  geom_point(size=3) +
  geom_smooth() + 
  labs(title = "Survival over Ub (final census)",
       x = "Bottomstress (Ub)",
       y = "% Survival (yield)") +
  theme_bw() +
  ylim(0,1) +
  xlim(0,.8) +
  theme(panel.border = element_rect(color = "black", fill = NA),
        axis.line = element_line(color = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14), 
        legend.text = element_text(size = 12)) +
  facet_wrap(~Reef, scales = "free")

plot10
ggsave("Fig10 - Survival final census Ub - yield.png", plot=plot10, width = 14, height = 6, dpi = 300)
```

### Fig11 - Point - survival final census over median current

```{r}
plot11 <- ggplot(merged_data_finalcensus, aes(x = median_speed, y = Perc_SurvDev, color = Reef)) +
  geom_point(size=3) +
  geom_smooth() + 
  labs(title = "Survival over median current (final census)",
       x = "median current (m/s)",
       y = "% Survival (yield)") +
  theme_bw() +
  ylim(0,1) +
  xlim(0,.3) +
  theme(panel.border = element_rect(color = "black", fill = NA),
        axis.line = element_line(color = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14), 
        legend.text = element_text(size = 12)) +
  facet_wrap(~Reef, scales = "free")

plot11
ggsave("Fig11 - Survival final census median current - yield.png", plot=plot11, width = 14, height = 6, dpi = 300)
```

```{r}
#####

Fig2 <- SummaryData |>
  ggplot(aes(y=PerSurv_t6*100, x=Current_Marotte_median))+
  geom_point()+
  geom_smooth(se=FALSE)+
  facet_grid(~Reef)+
  ylim(0,100)

Fig3 <- SummaryData %>%
  ggplot(aes(y=PerSurv_t6*100, x=NomWaveEnergy))+
  geom_point()+
  geom_smooth(se=FALSE)+
  facet_grid(~Reef)+
  ylim(0,100)+
  ylab("Percent survival")


ggsave("Fig1_ModelWaveEnergy_t6.png", Fig1, width = 6, units = "in")
ggsave("Fig2_MarotteaveEnergy_t6.png", Fig2, width = 6, units = "in")
ggsave("Fig3_NominalWaveEnergy_t6.png", Fig3, width = 6, units = "in")


Fig4 <- Data %>%
  ggplot(aes(y=PerSurv*100, x=ModelWaveEnergy))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  facet_grid(~Reef)+
  ylim(0,100)

Fig5 <- Data %>%
  ggplot(aes(y=PerSurv*100, x=NomWaveEnergy))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  facet_grid(~Reef)+
  ylim(0,100)
  
Fig6 <- Data %>%
  ggplot(aes(y=PerSurv*100, x=ModelWaveEnergyExt))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  facet_grid(~Reef)+
  ylim(0,100)+
  ylab("Percent survival")


ggsave("Fig4_ModelWaveEnergy_lm_t6.png", Fig4, width = 6, units = "in")
ggsave("Fig5_NomWaveEnergy_lm_t6.png", Fig5, width = 6, units = "in")
ggsave("Fig6_ModelWaveEnergyExt_lm_t6.png", Fig6, width = 6, units = "in")

Data %>%
  filter(Reef != "Moore") %>%
  ggplot(aes(y=PerSurv*100, x=SedPodAvg))+
  geom_point()+
  geom_smooth(se=FALSE)+
  facet_grid(~Reef)+
  ylab("Percent survival")+
  scale_x_continuous(trans = "log10")

Data %>%
  filter(Reef != "Moore") %>%
  ggplot(aes(y=PerSurv*100, x=TurfPodAvg))+
  geom_point()+
  geom_smooth(se=FALSE)+
  facet_grid(~Reef)+
  ylab("Percent survival")+
  scale_x_continuous(trans = "log10")

       
```
